Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-09-08 Experiments.tex
Address reviewers' comments: please review/iterate

We designed experiments to characterize the performance of the NGE on Titan,
with an emphasis on understanding its overhead and thus the cost of
introducing new functionalities. We perform three groups of experiments in
which we investigate the weak scalability, weak scalability with multiple
generation, and strong scalability of the NGE.

>>>>

We designed experiments to characterize the performance of the NGE on Titan,
with an emphasis on understanding its overhead and thus the cost of
introducing new functionalities. We perform three groups of experiments in
which we investigate the weak scalability, weak scalability with multiple
generation, and strong scalability of the NGE\@.

### 2017-09-08 PandaEvolution.tex
Address reviewers' comments: please review/iterate

\ifreview
Comments to address in this section:
\begin{enumerate}
	\color{red} 
	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
	is abruptly stopped by the description of Monte Carlo event generation
	and detector simulation (3A).
	\item REVIEWER 2: The single listing (3B) of similar projects within the
	collaboration does not detail whether this is derived or separate work.
	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
	message queues, sockets or other means? Are the restrictions a
	shortcomming of TITAN, or the majority of HPC environments? Given the
	proclaimed importance of HPC resources, a generic evaluation seems
	needed.
	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
	pilot model, performing both resource acquisition and workload execution?
	How do pilots conflict with backfill slots, when the former are designed
	to acquire resources independent of workload requirements? Why does each
	Broker instance query and use backfill separately? Given that this seems
	to be a major bottleneck, the choice of such a design begs for
	motivation.
	\item REVIEWER 2: Some facts about WLCG seem outdated:
	\begin{itemize}
		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
		[REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
		\item WLCG sites are claimed to provide only 8 cores per worker node
		(§IIIB). WLCG has a significant number of sites providing 16 or more
		cores per worker node [REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
	\end{itemize}
	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
	while their captions are not sufficient to understand them. The numbering
	is only partially explained in the text.
\end{enumerate}
\fi

The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

% Supercomputers are designed to support parallel computation that requires
% runtime communication. Jobs are executed across multiple nodes, each node
% calculating a small part of the problem and communicating with other node,
% typically via MPI\@. Accordingly, supercomputers have large number of
% worker nodes, connected through a high-speed, low-latency dedicated
% network. Each worker node has multicore CPUs, usually augmented with
% Graphics Processing Units (GPUs) or other specialized co-processors.

@@

% The Titan supercomputer, current number three on the Top 500
% list,%~\cite{top500}, 
% is a Cray XK7 system with 18,688 worker nodes and a total
% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
% Each worker node has an AMD Opteron 6274 16-core CPU,
% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
% inter-node MPI messaging. Titan is served by the Spider
% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
% Linux, a run time environment based on SUSE Linux Enterprise Server.
%
% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
% data transfer nodes (DTNs). Titan's authentication and authorization model
% is based on two-factor authentication with a RSA SecurID key. Login nodes
% and DTNs have out/inbound wide area network connectivity while worker nodes
% have only local network access. Fair-share and allocation policies are in
% place both for the PBS batch system and shared file systems.

@@

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the  Monte Carlo detector simulation task is most
suitable for execution on Titan at the moment. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and small
input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas}. %,de2013delphes,ritsch2014atlas}. 
The MC workflow consists of four main stages: event generation, detector
simulation, digitization, and reconstruction. Event generation creates sets
of particle four-momenta via different generators, e.g.,
PYTHIA, %~\cite{sjostrand2006pythia}, 
HERWIG, %~\cite{corcella2001herwig} 
and many others. Geant4 simulates the ATLAS detector and the interaction
between the detector and particles. Each interaction creates a so-called hit
and all hits are collected and passed on for digitalization, where hits are
further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level
objects like particles and jets.

% ---------------------------------------------------------------------------
\subsection{PanDA Broker}
\label{ssec:panda_titan}

@@

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request
as many worker nodes and walltime as a queue can offer. On the contrary, jobs
sized according to an available backfill slot depend on the number of worker
nodes and walltime that cannot be given to any other job at that moment in
time.

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given
the size of Titan this translates into a substantial capacity. Every year,
about 10\% of Titan's capacity remains unused~\cite{barker2016us},
corresponding to an average of 30,000 unused cores (excluding GPU cores).
This equals to roughly 10\% of the overall capacity of WLCG\@.

Given the communication requirements of PanDA Pilots and the unused capacity
of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
nodes of Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of
PanDA Pilot and its stand-alone architecture, this prototype called `PanDA
Broker' 
% implements functionalities to: (i) interrogate Titan about backfill
% availability via a dedicated Moab scheduler command; (ii) pull ATLAS jobs
% and events from PanDA Server; (iii) wrap the payload of ATLAS jobs into MPI
% scripts; (iv) submitting workload-specific MPI scripts to Titan's PBS batch
% system and monitor their execution; and (v) staging and preparing
% input/output files. Backfill querying, payload wrapping, and scripts
% submission required a new implementation while pulling ATLAS job and
% events, and file staging were inherited from PanDA Pilot.
%
% Backfill querying is performed via a dedicated Moab scheduler command while
% a tailored Python MPI script is used to execute the payload of ATLAS jobs.
% This enables the execution of unmodified Grid-centric, ATLAS jobs on Titan.
% Typically, a MPI script is workload-specific as it sets up the execution
% environment for a specific payload. This involves organization of worker
% directories, data management, optional input parameters modification, and
% cleanup on exit. Upon submission, a copy of the MPI script runs on every
% available worker node, starting the execution of the ATLAS job's payload in
% a subprocess and waits until its completion.
%
% MPI scripts are submitted to Titan's PBS batch system via
% RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the
% OGF GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for
% Grid Applications (SAGA) offers a unified interface to diverse job
% schedulers and file transferring services. In this way, SAGA provides an
% interoperability layer that lowers the complexity of using distributed
% infrastructures. Behind the API façade, RADICAL-SAGA implements a adaptor
% architecture: each adaptor interface the SAGA API with different middleware
% systems and services, including the PBS batch scheduler of Titan.
%
% The data staging capabilities of the PanDA Broker are implemented via a
% file system shared among DTNs and worker nodes. The input files with the
% events of the ATLAS jobs are downloaded on the shared filesystem from the
% data center of Brookhaven National Laboratory (BNL). The MPI script setup
% process includes making the location of these files available to the
% payload of the ATLAS's jobs. The PanDA Broker can locate the payload's
% output files on the shared filesystem and transfer them from Titan BNL\@.
%
% Once deployed on Titan, every PanDA Broker 
supports the execution of MC detector simulations in 9 steps.

Each PanDA Broker deployed on Titan queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
Broker pulls jobs' input files from BNL to the OLCF Lustre file system
(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
creates an MPI script, wrapping enough ATLAS jobs' payload to fit the
backfill slot. PanDA Broker submits the MPI script to the Titan's Torque
batch system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).

@@

While PanDA Broker implementation is resource specific, it was successfully
ported to other supercomputers, including the HPC2 at the National Research
Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
Edison/Cori at the National Energy Research Scientific Computing Center
(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
Centre (LRZ)~\cite{barreiro2016panda}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers
    indicates the execution process of a detector simulation job, part of the
    production ATLAS MC workflow.}
\label{fig:panda_broker}
\end{figure}
\vspace{-0.04in}

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
	\color{red} 
	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
	is abruptly stopped by the description of Monte Carlo event generation
	and detector simulation (3A).\mtnote{I read the section twice and I did
	not see such an abrupt transition.}
	\item REVIEWER 2: The single listing (3B) of similar projects within the
	collaboration does not detail whether this is derived or separate
	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
	within the existing ATLAS software infrastructure. Specified that the
	ATLAS team has ported PanDA Broker to other supercomputers.}
	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
	message queues, sockets or other means? Are the restrictions a
	shortcomming of TITAN, or the majority of HPC environments? Given the
	proclaimed importance of HPC resources, a generic evaluation seems
	needed.\mtnote{Added a sentence explaining that DTNs are not part of
	Titan's compute nodes.}
	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
	pilot model, performing both resource acquisition and workload
	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
	are developing NGE.} How do pilots conflict with backfill slots, when the
	former are designed to acquire resources independent of workload
	requirements?\mtnote{AFAIK, we never state that pilot conflict with
	backfill. We argue that the latter mitigates the absence of the former.}
	Why does each Broker instance query and use backfill separately? Given
	that this seems to be a major bottleneck, the choice of such a design
	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
	Broker supported rapid deployment and incremental optimization while
	incurring in functional limitations.}
	\item REVIEWER 2: Some facts about WLCG seem outdated:
	\begin{itemize}
		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
		[REBUSCPU].
		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
		\item WLCG sites are claimed to provide only 8 cores per worker node
		(§IIIB). WLCG has a significant number of sites providing 16 or more
		cores per worker node [REBUSCPU].
		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
		would require fairly deep rewriting on Section IVB. I edited the
		sentence referring to 8 cores eliminating the reference to compute
		nodes. Less than ideal but I did not find a better solution. Any
		suggestion?}
	\end{itemize}
	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
	while their captions are not sufficient to understand them. The numbering
	is only partially explained in the text.\mtnote{Added a reference to
	where the numbering of the figures are explained in the text. I
	double-checked and the numbering of both figure is fully explained in the
	text.}
\end{enumerate}
\fi

The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
600,000 cores at more than 100 Grid sites, with an aggregated performance of
8 petaFLOPS\@. This capacity will be sufficient for the planned analysis
and data processing, but it will be insufficient for the Monte Carlo
production workflow and any extra activity. To alleviate these challenges,
ATLAS is expanding the current computing model to include additional
resources such as the opportunistic use of supercomputers.

@@

@@

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the  Monte Carlo detector simulation task is most
suitable for execution on Titan at the moment. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and small
input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for \(\approx\)60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas}. The MC workflow consists of four main
stages: event generation, detector simulation, digitization, and
reconstruction. Event generation creates sets of particle four-momenta via
different generators, e.g., PYTHIA, HERWIG, and many others. Geant4 simulates
the ATLAS detector and the interaction between the detector and particles.
Each interaction creates a so-called hit and all hits are collected and
passed on for digitalization, where hits are further processed to mimic the
readout of the detector. Finally, reconstruction operates local pattern
recognition, creating high-level objects like particles and jets.

% ---------------------------------------------------------------------------
\subsection{PanDA Broker}\label{ssec:panda_titan}

@@

Compared to pilots, backfill has the disadvantage of limiting the amount of
resources that can be requested. Pilots are normal jobs: they can request as
many worker nodes and walltime as a queue can offer. On the contrary, jobs
sized according to an available backfill slot depend on the number of worker
nodes and walltime that cannot be given to any other job at that moment in
time.

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given
the size of Titan this translates into a substantial capacity. Every year,
about 10\% of Titan's capacity remains unused~\cite{barker2016us},
corresponding to an average of 30,000 unused cores (excluding GPU cores).
This equals to roughly 5\% of the overall capacity of WLCG\@.

Given the communication requirements of PanDA Pilots and the unused capacity
of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
nodes of Titan (Fig.~\ref{fig:panda_broker}). This prototype called `PanDA
Broker' maintains the core modules of PanDA Pilot and its stand-alone
architecture. This imposes functional trade-offs (e.g., single-threaded
architecture, single MPI PBS script submission) but allows for rapid adoption
and iterative optimization. PanDA Brokers are deployed on DTNs because these
nodes are part of the OLCF infrastructure and can access Titan without RSA
SecureID authentication. It should be noted that DTNs are not part of Titan's
worker nodes and, therefore, are not used to execute Titan's jobs.

Currently, up to 20 PanDA Brokers operate within the existing ATLAS
production software infrastructure, each supporting the execution of MC
detector simulations in 9 steps. Each broker queries the PanDA Server for
ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
Broker pulls jobs' input files from BNL Data Center to the OLCF Lustre file
system (Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab
scheduler about the current available backfill slot
(Fig.~\ref{fig:panda_broker}:3) and creates an MPI script, wrapping enough
ATLAS jobs' payload to fit the backfill slot. PanDA Broker submits the MPI
script to the Titan's PBS batch system via RADICAL-SAGA
(Fig.~\ref{fig:panda_broker}:4).

@@

PanDA Broker implementation is resource specific but the ATLAS team has
ported it to other supercomputers, including the HPC2 at the National
Research Center ``Kurchatov Institute''
(NRC-KI)~\cite{belyaev2015integration}, Edison/Cori at the National Energy
Research Scientific Computing Center (NERSC)~\cite{barreiro2016panda}, and
SuperMUC at the Leibniz Supercomputing Centre (LRZ)~\cite{barreiro2016panda}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers
    indicates the execution process of a detector simulation job described in
    \S\ref{ssec:panda_titan}.}\label{fig:panda_broker}
\end{figure}
\vspace{-0.04in}

### 2017-09-08 PandaOverview.tex
Comments to address before submission
Address reviewers' comments: please review/iterate

PanDA is a Workload Management System (WMS) %~\cite{marco2009glite} 
designed to support the execution of distributed workloads and workflows via
pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high
throughput execution of tasks via multi-level scheduling while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. 
% The design of PanDA WMS started in 2005 to support ATLAS\@. PanDA went into
% production for the LHC Run 1 on 2009, and was then extended with new
% subsystems for Run 2 in 2015.

% ---------------------------------------------------------------------------
% \subsection{Design}
% \label{ssec:panda_design}

% PanDA's application model assumes tasks grouped into workloads and
% workflows. Tasks represent a set of operations performed on data sets
% stored in one or more input files. Tasks are decomposed into jobs, where
% each job consists of the task's operations performed on a partition of the
% task's data set. Since 2005, a certain amount of parallelism has been
% progressively introduced for job execution~\cite{crooks2012multi} but, so
% far, HEP applications have not required a message passing interface (MPI).

% PanDA's usage model assumes multitenancy of resources and at least two
% types of HEP users: individual researchers and groups executing so called
% `production' workflows. PanDA's security model is based on separation
% between authentication, authorization and accounting for both single users
% and group of users. Both authentication and authorization are based on
% digital certificates and on the virtual organization
% abstraction~\cite{foster2001anatomy}.

% Currently, PanDA's execution model is based on four main abstractions:
% task, job, queue, and pilot. Both tasks and jobs are assumed to have
% attributes and states and to be queued into a global queue for execution.
% Prioritization and binding of jobs are assumed to depend on the attributes
% of each job. Pilot is used to indicate the abstraction of resource
% capabilities. Each job is thought to be bound to one pilot and executed on
% the site where the pilot has been instantiated.

% In PanDA's data model, each datum refers to the recorded or simulated
% measurement of a physical process. Data can be packaged into files or other
% containers. As with jobs, data have both attributes and states, and some of
% the attributes are shared between events and jobs. Raw, reconstruction, and
% simulation data are assumed to be distributed across multiple storage
% facilities and managed by the ATLAS Distributed Data Management
% (DDM)%~\cite{garonne2012atlas}. 
% When necessary, datasets required by each job are assumed to be replicated
% over the network, both for input and output data.

% PanDA's design supports provenance and traceability for both jobs and data.
% Attributes enable provenance by linking jobs and data items, providing
% information like ownership or project affiliation. States enable
% traceability by providing information about the stage of the execution in
% which each job or data item is or has been.

% ---------------------------------------------------------------------------
% \subsection{Implementation and Execution}
% \label{ssec:panda_arch}

The implementation of PanDA WMS consists of several interconnected
subsystems,
% , most of them built from off-the-shelf and Open Source components.
% Subsystems communicate 
communicating via dedicated API or HTTP messaging, and implemented by one or
more modules. Databases are used to store eventful entities like tasks, jobs
and input/output data, and to store information about sites, resources,
logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
ATLAS workflows % (e.g., PanDA Event Service~\cite{calafiura2015atlas}), 
but we do not discuss them as they are not relevant to an understanding of
how PanDA has been ported to supercomputers. For a full list of subsystems
see Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a
diagrammatic representation of PanDA main subsystems, highlighting the
execution process of tasks while omitting monitoring details to improve
readability.

% During LHC Run 1, PanDA required users to perform a static conversion
% between tasks and jobs: tasks were described as a set of jobs and then
% submitted to the PanDA Server. This introduced inefficiency both with
% usability and resource utilization.%~\cite{borodin2015unified}. Ideally,
% users should conceive analyses in terms of one or more, possibly related
% tasks, while the execution manager (i.e., PanDA) should partition tasks
% into jobs, depending on execution constraints. Further, due to the static
% partitioning of tasks into jobs, PanDA instantiates pilots on sites without
% taking into account the heterogeneity and dynamicity of the resources. An
% optimal sizing of each job should take into account these properties.

% Another problem of static job sizing is that PanDA instantiates pilots on
% sites with different type of resources and different models of availability
% of those resources. An optimal sizing of each job should take into account
% these properties. For example, sites may offer cores with different speed,
% networking with different amount of bandwidth, and resources could be
% guaranteed to be available for a defined amount of time or could disappear
% at any point in time as it happens with opportunistic models of resource
% provision.

% JEDI was deployed for the LHC Run 2 to address these inefficiencies. 
Users submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that
stores them into a queue implemented by a database
(Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of different
size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA
Server for execution (Fig.~\ref{fig:architecture}:4).

@@

When a PanDA Pilot becomes available, it requests the Job Dispatcher module
of the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11).
The Job Dispatcher interrogates the Task Buffer module for a job that is
bound to the site of that pilot and ready to be executed. Task Buffer checks
the global queue (i.e., the PanDA database) and, upon availability, returns a
job to the Job Dispatcher. The Job Dispatcher dispatches that job to the
PanDA Pilot (Fig.~\ref{fig:architecture}:12).

A PanDA Pilot starts a monitoring process on receiving a job and forks a
subprocess to execute the job's payload. Input data are transferred
from the stage-in location (Fig.~\ref{fig:architecture}:13), the job's
payload is executed (Fig.~\ref{fig:architecture}:14) and once completed, output
is transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based
    execution process. Several subsystems, components, and architectural and
    communication details are abstracted to improve clarity.}
\label{fig:architecture}
\end{figure}

>>>>

\ifreview
Comments to address in this section:
\begin{itemize}
	\color{red} 
	\item REVIEWER 2: use of wrong wordings, such as "eventful" instead of
	the common term "stateful".\mtnote{Done.}
\end{itemize}
\fi

PanDA is a Workload Management System (WMS) %~\cite{marco2009glite} 
designed to support the execution of distributed workloads and workflows via
pilots~\cite{turilli2017comprehensive}. Pilot-capable WMS enable high
throughput execution of tasks via multi-level scheduling while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data.

The implementation of PanDA WMS consists of several interconnected
subsystems, communicating via dedicated API or HTTP messaging, and
implemented by one or more modules. Databases are used to store stateful
entities like tasks, jobs and input/output data, and to store information
about sites, resources, logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
ATLAS workflows but we do not discuss them as they are not relevant to an
understanding of how PanDA has been ported to supercomputers. For a full list
of subsystems see Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture}
shows a diagrammatic representation of PanDA main subsystems, highlighting
the execution process of tasks while omitting monitoring details to improve
readability.

Users submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that
stores them into a queue implemented by a database
(Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of different
size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA
Server for execution (Fig.~\ref{fig:architecture}:4).

@@

When a PanDA Pilot becomes available, it requests the Job Dispatcher module
of the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11).
The Job Dispatcher interrogates the Task Buffer module for a job that is
bound to the site of that pilot and ready to be executed. Task Buffer checks
the global queue (i.e., the PanDA DB) and, upon availability, returns a job
to the Job Dispatcher. The Job Dispatcher dispatches that job to the PanDA
Pilot (Fig.~\ref{fig:architecture}:12).

Each PanDA Pilot starts a monitoring process on receiving a job and forks a
subprocess to execute the job's payload. Input data are transferred from the
stage-in location (Fig.~\ref{fig:architecture}:13), the job's payload is
executed (Fig.~\ref{fig:architecture}:14) and once completed, output is
transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based
    execution process described in \S\ref{sec:panda_overview}. Several
    subsystems, components, and architectural and communication details are
    abstracted to improve clarity.}
\label{fig:architecture}
\end{figure}

### 2017-09-08 bckground.tex
Comments to address before submission
Address reviewers' comments: please review/iterate

% \jhanote{In this section we should talk less about PanDA and more about other
% WMS -- highlighting the fact that (i) there isn't a comprehensive systems
% (design, architecture and execution properties) discussion of other WMS so
% difficult to compare}\mtnote{Better?}

% \jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
% this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
% meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
% the pace at which other WMS were being developed, loosing trust in their ability
% to deliver production-grade capabilities. ATLAS USA decided to go solo and
% developed their own solution. PanDA proven to be able to deliver the
% usual pilot-based WMS capabilities at the right scale. For that, it was chosen
% over other solutions.}

% \jhanote{Also might this section be moved towards the rear, as this is  a
% practice paper and not a traditional research paper?}\mtnote{Done.}

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
design and architectural principles: centralization of task and resource
management, and of monitoring and accounting; distribution of task execution
across multiple sites; unification of the application interface; hiding of
resource heterogeneity; and collection of static and sometimes dynamic
information about resources.

@@

% -----------------------------------------------------------------------------
% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
% -----------------------------------------------------------------------------
% Currently, the implementation of PanDA WMS has at least two distinguishing
% features: the scale at which it operates distributed computing, and
% network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me
% know whether more unique (i.e., not present in other LHC WMS) features of
% PanDA implementation should be indicated}. PanDA concurrently supports
% distributed computations on up to 240,000 cores, submitting more than 365
% million jobs a year across 100 sites, and processing up to XXXXPB of data
% distributed worldwide on XX Tier 1-3 facilities. As for 2017, PanDA serves
% several thousand users, including XX production groups. The scale of this
% operation is unprecedented, making PanDA the WMS that manages the largest
% computational campaign in the world.

% PanDA was initially implemented to support computations across an unreliable
% and relatively expensive networking infrastructure. Accordingly, PanDA
% assumed a rigid model of data placement based on replication and long-term
% storage of data on the Grid sites. This model proven too wasteful, requiring
% storage space for unused and obsolete data. PanDA implementation was evolved
% to benefit from the development of the internet and of the networking
% infrastructure for scientific research. Today, PanDA constantly monitors
% network throughput and latency among its management components, Grid sites and
% data facilities. PanDA uses this information to schedule tasks on sites,
% depending on the estimated time required to download and/or replicate input
% data and to stage out output data.

% The dynamic management of data replication and download is an example of a
% generalized paradigm shift that distinguishes the implementation of PanDA from
% other WMS. New components are being implemented to support: (i) dynamic sizing
% of input dataset based on resource capabilities; (ii) dynamic sizing of the
% resources held by pilots both in terms of number of cores and duration of the
% pilot; and (iii) further abstraction of different types of resources to
% reconcile high-throughput and high-performance computing paradigms.
% -----------------------------------------------------------------------------

% -----------------------------------------------------------------------------
% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
% -----------------------------------------------------------------------------
% Each task may have an arbitrary number of properties like number of cores,
% executables, or input/output files. Tasks may have precedence interrelations,
% depending on their data dependences or any other type of dependency mandated
% by the application algorithms. Tasks with precedence relations have to be
% executed serially; otherwise, tasks can be executed concurrently. A workload
% is defined as a set of tasks that can be executed concurrently. As such, a
% workflow can be composed of a set of workloads.

% The terms ``task'' and ``job'' are also used inconsistently across communities
% that perform scientific computing. In this paper, ``job'' refers to the unit
% of work that is submitted to a local resource management system (LRMS), like
% the Slurm or PBS batch system of a cluster. As such, a task can become a job
% when is scheduled on a resource that exposes a LRMS but can also become a
% virtual machine or a container when bootstrapped on an infrastructure
% supporting virtualization. Tasks can be statically or dynamically grouped
% into jobs, depending on the resource capabilities and the task requirements.

% Usually, workflows are represented as graphs in which tasks are vertices and
% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but
% graphs with cycles have been used to represent workflows of
% workflows~\cite{}. In this paper, a workflow template is a type of graph
% while a workflow instance is a workflow template with specific vertices and
% edges. Further, a workflow instance is ``abstract'' when no resource
% properties are available for all vertices, ``concrete'' otherwise~\cite{}.
% When not qualified, the term ``workflow'' indicates a abstract workflow
% instance.

% Notes for refactoring:
% \begin{itemize}
%     \item Naming and nomenclature is a contribution
%     \item 1, 1.5 pages.
% \end{itemize}
% -----------------------------------------------------------------------------

>>>>

\ifreview
Comments to address in this section:
\begin{itemize}
	\color{red} 
	\item REVIEWER 2: the topic of other HPC usage models [OCCAM,FRHPC],
	elastic simulation jobs [ELASTICMC], opportunistic HPC usage [ATLASHPC]
	or proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.
\end{itemize}
\fi

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE\@; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS\@; and
PanDA~\cite{maeno2014evolution} for ATLAS\@. These systems implement similar
design and architectural principles: centralization of task and resource
management, and of monitoring and accounting; distribution of task execution
across multiple sites; unification of the application interface; hiding of
resource heterogeneity; and collection of static and sometimes dynamic
information about resources.

@@
